---
title: "Effect Size Calculations"
author: "Eleanor Tsai"
date: "2023-09-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("esc")
library(esc)
```

```{r Glass Delta 2 Function}
# This function calculates the Glass Delta 2 from a regression coefficient and the SD of the control group

calculate_glass_delta2 <- function(b, SD_control, N_treatment, N_control) {
  # Calculate the effect size (delta)
  delta <- b / SD_control
  
  # Calculate the variance of delta
  variance_delta <- (N_treatment + N_control) / (N_treatment * N_control) + (delta^2) / (2 * (N_control - 1))
  
  # Calculate the standard error of delta
  se_delta <- sqrt(variance_delta)
  
  # Calculate the lower and upper bounds of the confidence interval
  lower_bound <- delta - 1.96 * se_delta # 1.96 for 95% confidence interval
  upper_bound <- delta + 1.96 * se_delta
  
  # Return the delta and its confidence interval
  formatted_result <- cat(paste0(
    "Effect Size (Delta):", round(delta, 4),
    "\n95% Confidence Interval: [", round(lower_bound, 4), ",", round(upper_bound, 4), "]\n"
    )
  )
  
  return(formatted_result)
}
```

```{r SD(Y) Function}
# This function calculates the Glass Delta 2 from a regression coefficient and its standard error, by first approximating the SD of the full sample and using it in place of the SD of the control group
# Reference here (last formula on the page): https://handbook-5-1.cochrane.org/chapter_7/7_7_3_3_obtaining_standard_deviations_from_standard_errors.htm
# If CIs are given, the formula for the standard error is (upper_bound-lower_bound)/3.92

calculate_sdy <- function(b, SE_diff, N_treatment, N_control) {
  # Calculate SD(Y) using the formula
  SD_Y <- SE_diff * sqrt((N_treatment * N_control) / (N_treatment + N_control))
  
  # Call the Glass' Delta 2 function
  result <- calculate_glass_delta2(b, SD_Y, N_treatment, N_control)
  
  # Return the result
  return(SD_Y)
  return(result)
}

```

# Example Usage for Language IDHC-B Score from Fernald & Hidrobo 2011
(See Table 2 and Table 3)
```{r}
b <- 2.43
SD_control <- 34.2
N_treatment <- 797
N_control <- 399
calculate_glass_delta2(b, SD_control, N_treatment, N_control)
```

```{r}
b <- 2.43
SE_diff <- 1.75
N_treatment <- 797
N_control <- 399
calculate_sdy(b, SE_diff, N_treatment, N_control)
```

```{r SE to CI Function}
get_ci <- function(std_b, SE) {
  # This function calculates the confidence interval for a standardized regression coefficient (i.e. effect size; reported in terms of Z-score) given its standard error
  
  lower_bound <- std_b - (1.96 * SE)
  upper_bound <- std_b + (1.96 * SE)
  
  confidence_interval <- c(round(lower_bound, 4), round(upper_bound, 4))
  
  return(confidence_interval)
}
```

# Weaver et al. 2024
## Cognitive
```{r}
std_b <- .09
SE <- .04
get_ci(std_b,SE)
```
## Gross Motor
```{r}
std_b <- .09
SE <- .04
get_ci(std_b,SE)
```
## Fine Motor
```{r}
std_b <- .10
SE <- .04
get_ci(std_b,SE)
```

# Weaver et al. 2023
## HAZ
```{r}
std_b <- .06
SE <- .05
get_ci(std_b,SE)
```
## WAZ
```{r}
std_b <- .04
SE <- .04
get_ci(std_b,SE)
```
# Carneiro et al. 2021
## HAZ
```{r}
std_b <- .135
SE <- .061
get_ci(std_b,SE)
```
## WAZ
```{r}
std_b <- .054
SE <- .056
get_ci(std_b,SE)
```
## WHZ
```{r}
std_b <- -.05
SE <- .056
get_ci(std_b,SE)
```
## Communication
Table A8
```{r}
std_b <- .043
SE <- .058
get_ci(std_b,SE)
```
## Gross Motor
Table A8
```{r}
std_b <- .082
SE <- .059
get_ci(std_b,SE)
```
## Personal-Social
Table A8
```{r}
std_b <- -.095
SE <- .064
get_ci(std_b,SE)
```

# Levere et al. 2016
## WAZ
```{r}
std_b <- .025
SE <- .063
get_ci(std_b,SE)
```

## HAZ
```{r}
std_b <- -.176
SE <- .107
get_ci(std_b,SE)
```

## WHZ
```{r}
std_b <- .171
SE <- .101
get_ci(std_b,SE)
```
## Communication
```{r}
std_b <- -.067
SE <- .072
get_ci(std_b,SE)
```
## Gross Motor
```{r}
std_b <- -.105
SE <- .075
get_ci(std_b,SE)
```
## Fine Motor
```{r}
std_b <- -.114
SE <- .065
get_ci(std_b,SE)
```
## Socioemotional
```{r}
std_b <- -.106
SE <- .069
get_ci(std_b,SE)
```
# Premand & Barry 2020
## HAZ
```{r}
std_b <- -.117
SE <- .064
get_ci(std_b,SE)
```

## WHZ
```{r}
std_b <- .057
SE <- .055
get_ci(std_b,SE)
```

## WAZ
```{r}
std_b <- -.053
SE <- .070
get_ci(std_b,SE)
```

## Cognitive
```{r}
std_b <- -.047
SE <- .067
get_ci(std_b,SE)
```

## Socioemotional
```{r}
std_b <- -.054
SE <- .069
get_ci(std_b,SE)
```

# Paxson & Schady 2010
## TVIP Score
```{r}
get_ci(.013,.089)
```
## Short-Term Memory
```{r}
get_ci(-.021,.098)
```

## Behavior Problems
```{r}
get_ci(.069,.090)
```

## Fine Motor
```{r}
get_ci(.105,.068)
```
# Macours et al. 2012
## Language
```{r}
get_ci(.139,.050)
```
## Short-Term Memory
```{r}
get_ci(.156,.044)
```

## Social-Personal
```{r}
get_ci(.130,.047)
```

## Gross Motor
```{r}
get_ci(-.006,.046)
```

## Fine Motor
```{r}
get_ci(.038,.063)
```

# Datta et al. 2021 Control
## Language Learning
```{r}
get_ci(.18,.07)
```

## Fine Motor
```{r}
get_ci(.09,.07)
```

## Social Skills
```{r}
get_ci(.03,.09)
```

# Datta et al. 2021 Arm 1
## Language Learning
```{r}
get_ci(.1,.07)
```

## Fine Motor
```{r}
get_ci(-.03,.08)
```

## Social Skills
```{r}
get_ci(.12,.09)
```


# Datta et al. 2021 Arm 2
## Language Learning
```{r}
get_ci(.09,.07)
```

## Fine Motor
```{r}
get_ci(-.05,.08)
```

## Social Skills
```{r}
get_ci(.11,.09)
```


# Datta et al. 2021 Arm 3
## Language Learning
```{r}
get_ci(.09,.08)
```

## Fine Motor
```{r}
get_ci(-.04,.08)
```

## Social Skills
```{r}
get_ci(.18,.08)
```


# Ozer et al. 2009
## Behavior Problems Index
```{r}
calculate_glass_delta2(.71,4.20,702,243)
```
# Wong et al. 2013
## Cognitive Skills
```{r}
get_ci(-.32,.17)
```

## Language Skills
```{r}
get_ci(.25,.24)
```

## Level of Self-Management
```{r}
get_ci(.1,.27)
```

## Motor Skills of Hands
```{r}
get_ci(-.02,.23)
```

## Overall Physical Capacity
```{r}
get_ci(.11,.19)
```

# Lopez-Boo & Creamer 2019 Education
## Communication
```{r}
get_ci(.141,.072)
```

## Gross Motor
```{r}
get_ci(.022,.080)
```

## Personal-Social
```{r}
get_ci(.076,.071)
```

## Problem-Solving
```{r}
get_ci(.082,.087)
```

# Lopez-Boo & Creamer 2019 Health
## Communication
```{r}
get_ci(.362,.120)
```

## Gross Motor
```{r}
get_ci(.127,.154)
```

## Personal-Social
```{r}
get_ci(.351,.118)
```

## Problem-Solving
```{r}
get_ci(.519,.161)
```


# Gaentzsch 2019
## PPVT
Table 7 Younger Cohort
```{r}
get_ci(-.232,.178)
```

## Maths
Table 7 Younger Cohort
```{r}
get_ci(-.611,.231)
```


# Sanchez et al. 2020 Younger Siblings
Table 2
## PPVT
```{r}
get_ci(.207,.116)
```

# Sanchez et al. 2020 Older Siblings
Table 2
## PPVT
```{r}
get_ci(-.022,.115)
```


# Bouguen & Dillon 2024 T1
## MELQO Cognitive Endline
```{r}
get_ci(-.045,.061)
```
## MELQO Language Endline
```{r}
get_ci(-.104,.087)
```
## MELQO Fine Motor Endline
```{r}
get_ci(.018,.104)
```
## CREDI Motor (Gross Motor & Fine Motor) Follow-Up
```{r}
get_ci(.021,.069)
```
## CREDI Socioemotional Follow-Up
```{r}
get_ci(0,.068)
```

# Bouguen & Dillon 2024 T2
## MELQO Cognitive Endline
```{r}
get_ci(.062,.077)
```
## MELQO Language Endline
```{r}
get_ci(-.063,.100)
```
## MELQO Fine Motor Endline
```{r}
get_ci(-.118,.114)
```
## CREDI Motor (Gross Motor & Fine Motor) Follow-Up
```{r}
get_ci(-.043,.076)
```
## CREDI Socioemotional Follow-Up
```{r}
get_ci(-.047,.074)
```


# Bouguen & Dillon 2024 T3
## MELQO Cognitive Endline
```{r}
get_ci(.010,.078)
```
## MELQO Language Endline
```{r}
get_ci(-.088,.088)
```
## MELQO Fine Motor Endline
```{r}
get_ci(.191,.096)
```
## CREDI Motor (Gross Motor & Fine Motor) Follow-Up
```{r}
get_ci(.135,.070)
```
## CREDI Socioemotional Follow-Up
```{r}
get_ci(.133,.070)
```


# Bjorvatn et al. 2023 T2
## Emergent Literacy
```{r}
get_ci(.06,.06)
```
## Emergent Numeracy
```{r}
get_ci(.08,.06)
```
## Socioemotional
```{r}
get_ci(.01,.07)
```
## Motor (Gross Motor & Fine Motor)
```{r}
get_ci(.11,.06)
```


# Bjorvatn et al. 2023 T3
## Emergent Literacy
```{r}
get_ci(.16,.06)
```
## Emergent Numeracy
```{r}
get_ci(.1,.06)
```
## Socioemotional
```{r}
get_ci(.04,.07)
```
## Motor (Gross Motor & Fine Motor)
```{r}
get_ci(.19,.06)
```
